Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2016-02-01 paper.tex
final version

\maketitle
\begin{abstract} 
Various unification schemes interpret the complex phenomenology of quasars and luminous active galactic nuclei (AGN) in terms of a simple picture involving a central black hole, an accretion disc and an associated outflow. Here, we continue our tests of this paradigm by comparing quasar spectra to synthetic spectra of biconical disc wind models, produced with our state-of-the-art Monte Carlo radiative transfer code. Previously, we have shown that we could produce synthetic spectra resembling those of observed broad absorption line (BAL) quasars, but only if the X-ray X-ray luminosity was limited to 10$^{43}$~erg~s$^{-1}$. 
 {\bf Here, we introduce a simple treatment of clumping, and find that a filling factor of $\sim0.01$ 
moderates the ionization state sufficently for BAL features to form in the 
rest-frame UV at more realistic X-ray luminosities.} Our fiducial model shows good agreement with AGN X-ray properties and the wind produces strong line emission in, e.g., \la\ and \civline\ at low inclinations. At high inclinations, the spectra possess prominent LoBAL features. Despite these successes, we cannot reproduce all emission lines seen in quasar spectra with the correct equivalent-width ratios, and we find an angular dependence of emission-line equivalent width despite the similarities in the observed emission line properties of BAL and non-BAL quasars. Overall, our work suggests that biconical winds can reproduce much of the qualitative behaviour expected from a unified model, but we cannot yet provide quantitative matches with quasar properties at all viewing angles. Whether disc winds can successfully unify quasars is therefore still an open question.
\end{abstract}

@@

Despite the importance of disc winds in shaping quasar and AGN spectra,  
much of the underlying outflow physics remains uncertain. 
Several driving mechanisms have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflow, and in AGN, unlike O stars, 
the presence of strong X-ray emission can overionize the wind, 
causing it to `fail'.
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
An additional or alternative solution is that the wind is clumped 
{\bf \citep[e.g.][]{junk1983,weymann1985,hamann2013}}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma 
without requiring excessively large total column densities 
and mass-loss rates.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
{\bf 
\cite{krolik1981} showed that BLR clouds would be short-lived, suggesting 
that a confining mechanism is required, such as
magnetic confinement \citep[e.g.][]{dekool1995}. Alternatively,
the `line deshadowing instability' (LDI) inherent in line-driven winds
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984}
may be responsible for smaller-scale clumping, as is observed in O-star winds
\citep[][and references therein]{fullerton2011}.
Indeed, simple models of clumping have successfully explained the electron scattering 
wings of emission lines formed in these line-driven flows 
\citep{hillier1991eswingsmodel}.}
Complex substructures are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to LDIs 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.  
In summary, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario. However, it is equally important to note that there 
is no consensus view of the size or location of the clumps in the outflow.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}),
which are reprocessed by `macro-atoms' with associated transition probabilities.
{\bf We use the Sobolev approximation \citep[e.g.][]{sobolev1957,sobolev1960,rybickihummer1978}
to compute the location and optical depth of line interactions}.
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15. In this scheme, the energy packets interact with either two-level 
`simple atoms' or full macro-atoms. 
This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple atoms (see M15). In this study,
only H and He are treated as macro-atoms, because the process is computationally
intensive, and we expect recombination to be particularly important
in determining their level populations and resultant line emission.
We are also especially interested in the contribution to 
AGN spectra of \LA.  H13 treated all atoms in a two-level approximation.

@@

We include  free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
{\bf Adiabatic cooling in the cell is calculated using the estimator
\begin{equation}
C_{A} = f_V k_B T_e V (\nabla \cdot v) \left(n_e + \sum\limits_{i=1}^{ions} N_i\right),
\end{equation}
where $V$ is the cell volume, $T_e$ is the electron temperature,
$n_e$ is the electron density, $N_j$ is the number density of ion $j$
and $\nabla \cdot v$ is the divergence of the 
velocity field at the centre of the cell. The sum is over all ions 
included in the simulation.
For consistency, $k$-packets then have a finite 
destruction probability of 
\begin{equation}
P_{k,A} = \frac{C_{A}}{C_{tot}},
\end{equation}
where $C_{tot}$ is the sum over all cooling rates in the cell.
}

@@

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. {\bf We emphasise that the purpose of this
purely kinematic wind model is to provide a simple tool for 
exploring different outflow geometries and kinematics.}
A schematic is shown in Fig.~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} to explain the phenomenonology
of quasars and BALQSOs. {\bf However, we do not include the initial
vertical component present in the Elvis (2000) model, which was
invoked as a qualitative explanation for narrow absorption lines}.

@@

In our previous modelling efforts, we assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
A clumpy outflow offers a possible solution to the so-called `over-ionization problem' in 
quasar and AGN outflows {\bf \citep[e.g.][]{junk1983,weymann1985,hamann2013}}. 
This is the main motivation for incorporating clumping into our model.

{\bf Deciding on how to implement clumping into our existing wind models was not straightforward.
First, and most importantly, the physical scale lengths and density contrasts in AGN outflows are not well-constrained from observations or theory.  As a result, while one could envision in principle, clouds with a variety of sizes and density contrasts varying perhaps as function of radius, there would have been very little guidance on how to set nominal values of the various parameters of such a model.}  
Second, there are significant computational difficulties associated with adequately resolving and realistically modelling a series of small scale, high density regions with a MCRT
{\bf -- or for that matter, a hydrodynamical --} code. 
Given the lack of knowledge about the actual type of clumping, we have adopted
a simple approximation used successfully in stellar wind modelling, known as 
{\em microclumping} \citep[e.g.][]{hamann1998,hilliermiller1999,hamann2008}.

The underlying assumption of microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. {\bf This approach 
allows one to treat clumps only in terms of their volume filling factor, $f_V$, 
instead of having to specify separately their size and density distributions.
In our model, $f_V$ is independent of position (see section~4.4).
The inter-clump medium is modeled as a vacuum,
although the outflow is still non-porous and axisymmetric.
This approach therefore assumes that the inter-clump medium
is unimportant in determing the output spectrum, which
we expect to be true only when density constrasts are large and
the inter-clump medium is both very ionized and of low emissivity and opacity.
}
The density of the clumps is multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
processes that are linear in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the square of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.
In our models, the temperature is not fixed, and is instead set by balancing heating and 
cooling in a given cell. In the presence of an X-ray source, this thermal balance is 
generally dominated by bound-free heating and line cooling. The main effect of including 
clumping in our modelling is that it moderates the ionization state due to the increased 
density. This allows an increase in the ionizing luminosity, amplifying the amount of
bound-free heating and also increasing the competing line cooling term
(thermal line emission).

{\bf 
The shortest length scale in a Sobolev MCRT treatment such as ours 
is normally the Sobolev length, given by
\begin{equation}
l_S = \frac{v_{th}}{| dv/ds |}
\end{equation}
This is typically $\sim10^{13}$~cm near the disc plane, increasing outwards.
We use the mean density to calculate the Sobolev optical depth, which assumes that
$l_S$ is greater than the typical clump size.
Thus for the microclumping assumption to be formally correct, 
clumps should be no larger than $\sim10^{12}$~cm.
This size scale is not unreasonable for quasar outflows, as
\cite{dekool1995} suggest that BAL flows may have low filling factors with
clump sizes of $\sim10^{11}$~cm. }
% An additional concern with this approach to clumping is that the clumps
% may rapidly cool and expand into the inter-clump medium. This is a well-known
% issue in BLR and outflow models (REFs), and requires some kind of confining mechanism,
% such as magnetic confinement \cite[e.g.][]{dekool1995}. Alternatively, as already discussed,
% the LDI may cause clumpy flows as expected in stellar winds.}

@@

{\bf The $89^\circ$ viewing angle cannot represent a typical quasar within a unified model, as
extreme inclinations should be extremely underrepresented in quasar samples.
This is in part due to the compton-thick absorber or `torus' 
expected at high inclinations \cite[e.g.][]{antonucci1985,martinez2007}.
However, this spectrum does show that a wind model can
naturally reproduce quasar emission lines if the emissivity of the wind is 
increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.}

@@

Having selected an individual fiducial model from the simulation grid, it is important
to briefly explore how specialised this model is, and how small parameter
changes can affect the synthetic spectra. {\bf Fig.~\ref{fig:grid}
shows the EW for} \civline\ {\bf and} \mgline\ {\bf at a low inclination, 
and $BI$ for} \civline\ {\bf at a high inclination for the simulation
grid. We find that almost all the models with $f_V=0.1$ are over-ionized, and 
fail to produce strong} \civ\ {\bf BALs or emission lines. However, 
the models with $f_V=0.01$ generally produce} \civ\ {\bf BALs and emission lines.
The fiducial model is representative of this family of models and 
the spectra are generally similar, although it is clear from the figure
that we have selected a model near the upper end of the EW distributions.

We find that it is difficult to significantly increase line emission while
keeping the luminosity and mass loss rate of the system fixed.}
We show an additional point on figure 7 corresponding to a model with an order of
magnitude higher X-ray luminosity and double the mass loss rate. As expected, 
this results in far higher line EWs, but fails to produce BALs because
the collisionally excited emission swamps the BAL profile. In addition,
this model would lie well above the expected $L_{2kev}-L_{2500}$ 
relation in figure 5. Such a high X-ray luminosity could therefore 
not be the cause of the strong line emission seen in {\em all} Type 1 quasars.

@@

The work of JHM, SWM, NSH and CK is supported by the
Science and Technology Facilities Council (STFC),
via two studentships and a consolidated grant, respectively.
CK also acknowledges a Leverhulme fellowship.
{\bf We would like to thank the anonymous referee for a helpful
and constructive report.}
We would like to thank Omer Blaes, Ivan Hubeny and Shane Davis for their
assistance with \agn. We are grateful to Mike Brotherton, Mike DiPompeo,
Sebastien Hoenig and Frederic Marin for helpful correspondence regarding
polarisation measurements and orientation indicators.
We would also like to thank Daniel Proga, Daniel Capellupo, Sam Connolly and
Dirk Grupe for useful discussions.  Simulations were conducted using \py\ version 80,
and made use of the IRIDIS High Performance Computing Facility at the
University of Southampton. Figures were produced using the {\tt matplotlib} 
plotting library \citep{matplotlib}. This work made use of the Sloan Digital Sky Survey.
Funding for the Sloan Digital Sky Survey has been provided by
the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of
Science, and the Participating Institutions.

%% \texttt{mn2e.cls} \textsc{Latex} document class.

\bibliography{bibs/mybib.bib,bibs/stellar.bib,bibs/h14.bib,bibs/h13.bib,bibs/hamann.bib,bibs/krolik.bib}

>>>>

\maketitle
\begin{abstract} 
Various unification schemes interpret the complex phenomenology of quasars and luminous active galactic nuclei (AGN) in terms of a simple picture involving a central black hole, an accretion disc and an associated outflow. Here, we continue our tests of this paradigm by comparing quasar spectra to synthetic spectra of biconical disc wind models, produced with our state-of-the-art Monte Carlo radiative transfer code. Previously, we have shown that we could produce synthetic spectra resembling those of observed broad absorption line (BAL) quasars, but only if the X-ray X-ray luminosity was limited to 10$^{43}$~erg~s$^{-1}$. 
Here, we introduce a simple treatment of clumping, and find that a filling factor of $\sim0.01$ 
moderates the ionization state sufficently for BAL features to form in the 
rest-frame UV at more realistic X-ray luminosities. Our fiducial model shows good agreement with AGN X-ray properties and the wind produces strong line emission in, e.g., \la\ and \civline\ at low inclinations. At high inclinations, the spectra possess prominent LoBAL features. Despite these successes, we cannot reproduce all emission lines seen in quasar spectra with the correct equivalent-width ratios, and we find an angular dependence of emission-line equivalent width despite the similarities in the observed emission line properties of BAL and non-BAL quasars. Overall, our work suggests that biconical winds can reproduce much of the qualitative behaviour expected from a unified model, but we cannot yet provide quantitative matches with quasar properties at all viewing angles. Whether disc winds can successfully unify quasars is therefore still an open question.
\end{abstract}

@@

Despite the importance of disc winds in shaping quasar and AGN spectra,  
much of the underlying outflow physics remains uncertain. 
Several driving mechanisms have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflow, and in AGN, unlike O stars, 
the presence of strong X-ray emission can overionize the wind, 
causing it to `fail'.
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
An additional or alternative solution is that the wind is clumped 
\citep[e.g.][]{junk1983,weymann1985,hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma 
without requiring excessively large total column densities 
and mass-loss rates.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
\cite{krolik1981} showed that BLR clouds would be short-lived, suggesting 
that a confining mechanism is required, such as
magnetic confinement \citep[e.g.][]{dekool1995}. Alternatively,
the `line deshadowing instability' (LDI) inherent in line-driven winds
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984}
may be responsible for smaller-scale clumping, as is observed in O-star winds
\citep[][and references therein]{fullerton2011}.
Indeed, simple models of clumping have successfully explained the electron scattering 
wings of emission lines formed in these line-driven flows 
\citep{hillier1991eswingsmodel}.
Complex substructures are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to LDIs 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.  
In summary, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario. However, it is equally important to note that there 
is no consensus view of the size or location of the clumps in the outflow.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}),
which are reprocessed by `macro-atoms' with associated transition probabilities.
We use the Sobolev approximation \citep[e.g.][]{sobolev1957,sobolev1960,rybickihummer1978}
to compute the location and optical depth of line interactions.
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15. In this scheme, the energy packets interact with either two-level 
`simple atoms' or full macro-atoms. 
This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple atoms (see M15). In this study,
only H and He are treated as macro-atoms, because the process is computationally
intensive, and we expect recombination to be particularly important
in determining their level populations and resultant line emission.
We are also especially interested in the contribution to 
AGN spectra of \LA.  H13 treated all atoms in a two-level approximation.

@@

We include  free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling in the cell is calculated using the estimator
\begin{equation}
C_{A} = f_V k_B T_e V (\nabla \cdot v) \left(n_e + \sum\limits_{i=1}^{ions} N_i\right),
\end{equation}
where $V$ is the cell volume, $T_e$ is the electron temperature,
$n_e$ is the electron density, $N_j$ is the number density of ion $j$
and $\nabla \cdot v$ is the divergence of the 
velocity field at the centre of the cell. The sum is over all ions 
included in the simulation.
For consistency, $k$-packets then have a finite 
destruction probability of 
\begin{equation}
P_{k,A} = \frac{C_{A}}{C_{tot}},
\end{equation}
where $C_{tot}$ is the sum over all cooling rates in the cell.

@@

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. The purpose of this
purely kinematic wind model is to provide a simple tool for 
exploring different outflow geometries.
A schematic is shown in Fig.~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} to explain the phenomenonology
of quasars and BALQSOs. However, we do not include the initial
vertical component present in the Elvis (2000) model, which was
invoked as a qualitative explanation for narrow absorption lines.

@@

In our previous modelling efforts, we assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
A clumpy outflow offers a possible solution to the so-called `over-ionization problem' in 
quasar and AGN outflows \citep[e.g.][]{junk1983,weymann1985,hamann2013}. 
This is the main motivation for incorporating clumping into our model.

Deciding on how to implement clumping into our existing wind models was not straightforward.
First, and most importantly, the physical scale lengths and density contrasts in AGN outflows are not well-constrained from observations or theory.  As a result, while one could envision in principle, clouds with a variety of sizes and density contrasts varying perhaps as function of radius, there would have been very little guidance on how to set nominal values of the various parameters of such a model.
Second, there are significant computational difficulties associated with adequately resolving and realistically modelling a series of small scale, high density regions with a MCRT
-- or for that matter, a hydrodynamical -- code. 
Given the lack of knowledge about the actual type of clumping, we have adopted
a simple approximation used successfully in stellar wind modelling, known as 
{\em microclumping} \citep[e.g.][]{hamann1998,hilliermiller1999,hamann2008}.

The underlying assumption of microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to treat clumps only in terms of their volume filling factor, $f_V$, 
instead of having to specify separately their size and density distributions.
In our model, $f_V$ is independent of position (see section~4.4).
The inter-clump medium is modeled as a vacuum,
although the outflow is still non-porous and axisymmetric.
This approach therefore assumes that the inter-clump medium
is unimportant in determing the output spectrum, which
we expect to be true only when density constrasts are large and
the inter-clump medium is both very ionized and of low emissivity and opacity.
The density of the clumps is multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
processes that are linear in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the square of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.
In our models, the temperature is not fixed, and is instead set by balancing heating and 
cooling in a given cell. In the presence of an X-ray source, this thermal balance is 
generally dominated by bound-free heating and line cooling. The main effect of including 
clumping in our modelling is that it moderates the ionization state due to the increased 
density. This allows an increase in the ionizing luminosity, amplifying the amount of
bound-free heating and also increasing the competing line cooling term
(thermal line emission).

The shortest length scale in a Sobolev MCRT treatment such as ours 
is normally the Sobolev length, given by
\begin{equation}
l_S = \frac{v_{th}}{| dv/ds |}
\end{equation}
This is typically $\sim10^{13}$~cm near the disc plane, increasing outwards.
We use the mean density to calculate the Sobolev optical depth, which assumes that
$l_S$ is greater than the typical clump size.
Thus for the microclumping assumption to be formally correct, 
clumps should be no larger than $\sim10^{12}$~cm.
This size scale is not unreasonable for quasar outflows, as
\cite{dekool1995} suggest that BAL flows may have low filling factors with
clump sizes of $\sim10^{11}$~cm.
% An additional concern with this approach to clumping is that the clumps
% may rapidly cool and expand into the inter-clump medium. This is a well-known
% issue in BLR and outflow models (REFs), and requires some kind of confining mechanism,
% such as magnetic confinement \cite[e.g.][]{dekool1995}. Alternatively, as already discussed,
% the LDI may cause clumpy flows as expected in stellar winds.}

@@

The $89^\circ$ viewing angle cannot represent a typical quasar within a unified model, as
extreme inclinations should be extremely underrepresented in quasar samples.
This is in part due to the compton-thick absorber or `torus' 
expected at high inclinations \cite[e.g.][]{antonucci1985,martinez2007}.
However, this spectrum does show that a wind model can
naturally reproduce quasar emission lines if the emissivity of the wind is 
increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.

@@

Having selected an individual fiducial model from the simulation grid, it is important
to briefly explore how specialised this model is, and how small parameter
changes can affect the synthetic spectra. Fig.~\ref{fig:grid}
shows the EW for \civline\ and \mgline\ at a low inclination, 
and $BI$ for \civline\ at a high inclination for the simulation
grid. We find that almost all the models with $f_V=0.1$ are over-ionized, and 
fail to produce strong \civ\ BALs or emission lines. However, 
the models with $f_V=0.01$ generally produce \civ\ BALs and emission lines.
The fiducial model is representative of this family of models and 
the spectra are generally similar, although it is clear from the figure
that we have selected a model near the upper end of the EW distributions.

We find that it is difficult to significantly increase line emission while
keeping the luminosity and mass loss rate of the system fixed.
We show an additional point on figure 7 corresponding to a model with an order of
magnitude higher X-ray luminosity and double the mass loss rate. As expected, 
this results in far higher line EWs, but fails to produce BALs because
the collisionally excited emission swamps the BAL profile. In addition,
this model would lie well above the expected $L_{2kev}-L_{2500}$ 
relation in figure 5. Such a high X-ray luminosity could therefore 
not be the cause of the strong line emission seen in {\em all} Type 1 quasars.

@@

The work of JHM, SWM, NSH and CK is supported by the
Science and Technology Facilities Council (STFC),
via two studentships and a consolidated grant, respectively.
CK also acknowledges a Leverhulme fellowship.
We would like to thank the anonymous referee for a helpful
and constructive report.
We would like to thank Omer Blaes, Ivan Hubeny and Shane Davis for their
assistance with \agn. We are grateful to Mike Brotherton, Mike DiPompeo,
Sebastien Hoenig and Frederic Marin for helpful correspondence regarding
polarisation measurements and orientation indicators.
We would also like to thank Daniel Proga, Daniel Capellupo, Sam Connolly and
Dirk Grupe for useful discussions.  Simulations were conducted using \py\ version 80,
and made use of the IRIDIS High Performance Computing Facility at the
University of Southampton. Figures were produced using the {\tt matplotlib} 
plotting library \citep{matplotlib}. This work made use of the Sloan Digital Sky Survey.
Funding for the Sloan Digital Sky Survey has been provided by
the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of
Science, and the Participating Institutions.

%% \texttt{mn2e.cls} \textsc{Latex} document class. 
\setlength{\bibsep}{0pt}
\bibliography{bibs/mybib.bib,bibs/stellar.bib,bibs/h14.bib,bibs/h13.bib,bibs/hamann.bib,bibs/krolik.bib}

### 2016-02-01 paperbold.tex
final version

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. {\bf We emphasise that the purpose of this
purely kinematic wind model is to provide a simple tool for 
exploring different outflow geometries and kinematics, and
the model itself is not hydrodynamic.}
A schematic is shown in Fig.~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} to explain the phenomenonology
of quasars and BALQSOs. {\bf However, we do not include the initial
vertical component present in the Elvis (2000) model, which was
invoked as a qualitative explanation for narrow absorption lines}.

>>>>

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. {\bf The purpose of this
purely kinematic wind model is to provide a simple tool for 
exploring different outflow geometries.}
A schematic is shown in Fig.~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} to explain the phenomenonology
of quasars and BALQSOs. {\bf However, we do not include the initial
vertical component present in the Elvis (2000) model, which was
invoked as a qualitative explanation for narrow absorption lines}.
